Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-05-14 paper.qmd
Started updating the equations

Because we are working with distances between points, starting with data having a standard scale, e.g. [0, 1], is recommended. The default should take the aspect ratio produced by the NLDR $(r_1, r_2, ..., r_k)$ into account. When $k=2$, as in hexagon binning, the default range is $[0, y_{i,\text{max}}], i=1,2$, where $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}$ (@fig-NLDR-two-curvy). If the NLDR aspect ratio is ignored then set $y_ {2,\text{max}} = 1$. <!-- \times \frac{2}{\sqrt{3}}$. (The $\frac{2}{\sqrt{3}}$ accounts for the different height ($a_1$) and width ($a_2$) of a regular hexagon.) The scaling of data should take the size of the hexagons into account, but choice of number of bins should. -->

@@

The value for $b_2$ is computed by fixing $b_1$. Considering the upper bound of the first NLDR component, $a_1 > \frac{1+2q}{b_1 -1}$. Similarly, for the second NLDR component, $a_2 > \frac{r_2 + q(1 + r_2)}{(b_2 - 1)}$. Since $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons, $a_1 > \frac{2[r_2 + q(1 + r_2)]}{\sqrt{3}(b_2 - 1)}$. This is a linear optimization problem. Therefore, the optimal solution must occur on a vertex. Therefore, $b_2 = \Big\lceil1 +\frac{2[r_2 + q(1 + r_2)](b_1 - 1)}{\sqrt{3}(1 + 2q)}\Big\rceil$.

>>>>

Because we are working with distances between points, starting with data having a standard scale, e.g. [0, 1], is recommended. The default should take the aspect ratio produced by the NLDR $(r_1, r_2, ..., r_k)$ into account. When $k=2$, as in hexagon binning, the default range is $[0, y_{i,\text{max}}], i=1,2$, where $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = r_2/r_1$ (@fig-NLDR-two-curvy). If the NLDR aspect ratio is ignored then set $y_ {2,\text{max}} = 1$. <!-- \times \frac{2}{\sqrt{3}}$. (The $\frac{2}{\sqrt{3}}$ accounts for the different height ($a_1$) and width ($a_2$) of a regular hexagon.) The scaling of data should take the size of the hexagons into account, but choice of number of bins should. -->

@@

The value for $b_2$ is computed by fixing $b_1$. Considering the upper bound of the first NLDR component, $a_1 > (1+2q)/(b_1 -1)$. Similarly, for the second NLDR component, $a_2 > (r_2 + q(1 + r_2))/(b_2 - 1)$. Since $a_2 = \sqrt(3)a_1/2$ for regular hexagons, $a_1 > \frac{2[r_2 + q(1 + r_2)]}{\sqrt{3}(b_2 - 1)}$. This is a linear optimization problem. Therefore, the optimal solution must occur on a vertex. Therefore, $b_2 = \Big\lceil1 +\frac{2[r_2 + q(1 + r_2)](b_1 - 1)}{\sqrt{3}(1 + 2q)}\Big\rceil$.
